Optimal analysis of Discrete-time Time-Invariant Affine Systems

### 2018-10-01 paper.tex
new

\begin{document}
\maketitle
\begin{abstract}

\end{abstract}

@@

This interest for static analysis techniques comes from its exhaustivity. Indeed static analysis offers a deeper analysis compared to traditional test techniques which cannot cover all possible situations. The principle of static analysis consists in proving properties automatically on programs without executing them. Static analyzers only use the structures of programs not the values generated by the program.  Numerous various techniques exist but none of them cannot be complete. This means that no program can validate or disprove a given property for all programs. Static analysis is not only useful for programs but can also be applied to validate properties on dynamical systems in general. The principle is still the same : to prove properties on dynamical systems without any simulations.

@@

The main motivation of the paper is to propose a technique that can be, for a given discrete-time affine system and a given property, to prove or disprove the property on the system.

%To do this, we reformulate the verification problem as an optimization problem with dynamical system contraints. This optimization problem is then solved exactly. The exact optimal solution can guarantee whether the property holds or not. Note that this particular subclass of optimization problem is called robust-to-dynamics optimization in~\cite{7403149}.   
\paragraph{Contributions}
Compute an integer that permits to solve a optimization problem with reachability constraints.
This problem is then solved exactly
From the computer science point of view, the motivation, we can prove or disprove a property

\paragraph{Related works .}
For the static analysis of discrete-time affine systems, numerous works in computer science or control theory exist. 
in computer science as we said earlier, works are decomposed into two world model checking community and abstract interpretation.

Model-checking approach as said earlier can disprove a property by the synthesis of counter-examples. The direct validation of the property would require an infinite number of runs.

@@

Our very first concern is the resolution of the verification problem for the class of discrete-time affine dynamical systems. This problem is widely studied in computer science community~\cite{mine} or in control theory~\cite{safety}. Optimization theory 
~\cite{optim} are more and more used to solve verification problems. The proposed approach follows this line of research : our verification problem is reduced to solve an optimization problem with dynamical system constraints. Actually this particular subclass of optimization problems is called the robust-to-dynamics optimization~\cite{ahmadi}.

In our presentation, we consider the maximization of a convex quadratic function over the reachable values set of an asymptotic stable affine system. For verification purposes, we assume that the initial condition set is a polytope. The difficulty in solving the optimization problem is to handle infinite sequences in the constraints set. To solve this optimization problem, we truncate the infinite sequences belonging to reachable values set at a rank which is uniform with respect to the initial conditions i.e. valid for all possible initial conditions.   
This uniform rank is constructed from basic tools of matrix theory such as matrix norms, extremal eigenvalues or generalized Rayleigh quotients. We also need a solution of the discrete Lyapunov equation that permits to define a useful matrix norm. Finally, we introduce a new non-convex semi-definite program to compute the optimal uniform rank. The exact resolution of this semi-definite problem is, for the moment, left open and we just provide an over-approximation of its optimal value. By doing so, it does not degrade the exact resolution of the verification problem.

In practice, we use semi-definite programming to solve the discrete Lyapunov matrix equation and to compute extremal eigenvalues that we need. We end the talk by some experimental results that measure the accuracy of our approach.

Verification linear /affine systems
Optimization tools for verification
Formal methods tools 
Control problems with reachability

\paragraph{Organization of the paper}
Problem statement, the main results, the computation of the integer, the experimentation and finally a conclusion.
\paragraph{Notations}
In this paper, we denote by $\Idd$ the identity matrix of size $d\times d$. We denote, for a matrix $M$, by $M^\intercal$ the transpose of $M$. The set of the (square) symmetric matrices  (the matrices $M$ such that $M=M^\intercal$) of size $d$ is denoted by $\sym$.

@@

\section{Problem statement}
In this paper, we are interested in proving automatically some properties on a discrete-time affine system. We start by briefly recalling the notion of affine systems and their reachable values set. Then we present the optimization problem that we have to solve to prove or disprove a property on an affine system.

@@

The problem in the computation of the optimal value of Problem~\eqref{pbopt} resides, first, in the infinite number of quadratic optimization problems that we have to solve. In a second time, the quadratic optimization problems are non-necessarily concave and thus can be difficult to solve. 
\section{Results}

This section contains the main contribution of the paper i.e. the computation of the optimal value of Problem~\eqref{pbopt}. This latter computation uses a discretization of initial values set and the computation of a finite integer after which we know that the optimal cannot be reached. In this section, we begin to recall some basic results that we need in our development. We then study the case of linear systems i.e. where $b=0$ in Eq.~\eqref{context}. Finally, we show how to use the linear case to solve the case where the system is affine.   
\subsection{Basic notions}

@@

To complete the computations, we have to deal with $\lfrak$. We suppose that the integer $\ks$ is given and then to compute $\lfrak$ we employ the same technique used for $\mu(\cdot)$. It suffices to browse the vertices of $\xin$ once since we evaluate, for each vertice, $\min\{x^\intercal Q x, x^\intercal {A^{\ks}}^\intercal Q {A^{\ks}}x+q^\intercal A^{\ks} x\}$ and take the maximal value. 
\section{Experiments}
We decompose the experiments into two sub-classes: the linear and affine systems. Those sub-classes are separeted in three parts : the  case where the objective function is a quadratic homogeneous function; the linear case and the non-homogeneous quadratic objective function. For each cut, we provide a numerical example.  
\subsection{Linear systems}

@@

\subsubsection{Random matrices}
\section{Conclusion and Future Works} 
In this paper, we introduce a new class of optimization problems. This class solves verification problem coming from computer science. This class can be described as follows : the objective function is convex and quadratic and the decision variable is constrained to belong to the reachable values set of a discrete-time affine system which is not a convex set. The main issue is the infinite horizon problem. To maximize the function we can wait for an arbitrary long time. Thus the paper proposes to compute, in the case of stable affine system, a finite horizon after which the search of an optimal value is useless. The problem becomes a finite horizon problem which can be solved in a finite time.

>>>>

\begin{document}
\maketitle
\begin{abstract}
 Our very first concern is the resolution of the verification problem for the class of discrete-time affine dynamical systems. This verification problem is turned into an optimization problem where the constraint set is the reachable values set of the dynamical system. To solve this optimization problem, we truncate the infinite sequences belonging to reachable values set at some step which is uniform with respect to the initial conditions. In theory, the best possible uniform step is the optimal solution of a non-convex semi-definite program. In practice,  we propose a methodology to compute an uniform step that over-approximate the best solution.
\end{abstract}

@@

This interest for static analysis techniques comes from the cover of them. Indeed static analysis offers a deeper analysis compared to traditional test techniques which cannot cover all possible situations. The principle of static analysis consists in proving properties automatically on programs without executing them. Static analyzers only use the structures of programs not the values generated by the program.  Numerous various techniques exist but none of them cannot be complete. This means that no program can validate or disprove a given property for all programs. Static analysis is not only useful for programs but can also be applied to validate properties on dynamical systems in general. The principle is still the same : to prove properties on dynamical systems without any simulations.

@@

The main motivation of the paper is to propose a technique that can be, for a given discrete-time affine system and a given property, to prove or disprove the property on the system. 
%This problem is widely studied in computer science community~\cite{mine} or in control theory~\cite{safety}. Optimization theory 
%~\cite{optim} are more and more used to solve verification problems. The proposed approach follows this line of research : our verification problem is reduced to solve an optimization problem with dynamical system constraints. Actually this particular subclass of optimization problems is called the robust-to-dynamics optimization~\cite{ahmadi}.

%To do this, we reformulate the verification problem as an optimization problem with dynamical system contraints. This optimization problem is then solved exactly. The exact optimal solution can guarantee whether the property holds or not. Note that this particular subclass of optimization problem is called robust-to-dynamics optimization in~\cite{7403149}.

\paragraph{Related works .}
%For the static analysis of discrete-time affine systems, numerous works in computer science or control theory exist. 
%in computer science as we said earlier, works are decomposed into two world model checking community and abstract interpretation. 
%
%Model-checking approach as said earlier can disprove a property by the synthesis of counter-examples. The direct validation of the property would require an infinite number of runs.  
The paper solves a formal verification problem. Numerous works (for example~\cite{DBLP:journals/corr/abs-1111-5223,gonnord2014abstract,mine-ESOP16,Sankaranarayanan_BenSassi__2017__Template}) about the resolution of the problem by abstract interpretation are available. They are mainly concerned in the representation of the reachable values. The over-approximations produced are not sufficient to solve a specific verification problem with a given property.

Hence, we only consider related works which produce a proof for a given property. Moreover, we are focused on close techniques based on optimization theory and or Lyapunov functions to solve the verification problem.

@@

\paragraph{Contributions}
Our very first concern is the resolution of the verification problem for the class of discrete-time affine dynamical systems. We assume that the given property is written as the sublevel set of some quadratic function. 
We reduced the verification problem to an optimization problem of the form :
\begin{equation}
\label{eqintro}
\sup\{f(x)\mid x_{k+1}=g(x_k),\ x_0\in\xin\}
\end{equation}
where $f$ is a quadratic function, $g$ an affine function and $\xin$ is a polytope.
Note that this optimization formulation has been already used in previous papers of the author for verification problems.

The optimal value~\eqref{eqintro} only depends on the initial values i.e. is equal to :
\[
\sup\{f( g^k (x))\mid  x_0\in\xin,\ k\in\nn\}\enspace.
\]
In the paper, we compute a number of steps $\bigk$ after which the search of the optimal value is useless i.e.
 \begin{equation}
\label{eqintro2}
\sup\{f( g^k (x))\mid  x_0\in\xin,\ k\in\nn\}=
\sup\{f( g^k (x))\mid  x_0\in\xin,\ k\in \{0,\ldots,\bigk\}\}
\end{equation}
This number of steps $\bigk$ is valid for all possible initial values. We are looking for the best possible integer $\bigk$ for which 
Equation~\eqref{eqintro2} is true. We propose to construct a family af integers parameterized by some positive reals and matrices solution of a discrete Lyapunov equation. The approach proposed in the paper to construct the family of integeres uses basic tools from matrix theory such as the discrete Lyapunov matrix equation, minimal and maximal eigenvalues, matrix norms and generalized Raleigh quotients. In theory, the best integer that we can propose is then the optimal value of some non-convex semi-definite program.  
The exact resolution of this semi-definite problem is, for the moment, left open and we just provide an over-approximation of its optimal value. By doing so, it does not degrade the exact resolution of the verification problem.   However, for the moment, we propose in practice the computation of two integeres which over-approximate the best value. 
%In practice, we use semi-definite programming to solve the discrete Lyapunov matrix equation and to compute extremal eigenvalues that we need. We end the talk by some experimental results that measure the accuracy of our approach.
%
%Verification linear /affine systems
%Optimization tools for verification
%Formal methods tools 
%Control problems with reachability

\paragraph{Organization of the paper}
Section~\ref{statement} recalls in details the problem statement that is the formulation of our verification problem and its translation into a verification problem. Section~\ref{mainresults} presents the main results. We give details about the construction of the family of integers solutions of Eq.~\eqref{eqintro2}. Section~\ref{compuations} describes the practical idea used in the implementation.  Section~\ref{experiments} is devoted to the experimentation. Finally, Section~\ref{conclusion} concludes and opens future researcjh directions.
\paragraph{Notations}
In this paper, we denote by $\Idd$ the identity matrix of size $d\times d$. We denote, for a matrix $M$, by $M^\intercal$ the transpose of $M$. The set of the (square) symmetric matrices  (the matrices $M$ such that $M=M^\intercal$) of size $d$ is denoted by $\sym$.

@@

\section{Problem statement}
\label{statement}
In this paper, we are interested in proving automatically some properties on a discrete-time affine system. We start by briefly recalling the notion of affine systems and their reachable values set. Then we present the optimization problem that we have to solve to prove or disprove a property on an affine system.

@@

The problem in the computation of the optimal value of Problem~\eqref{pbopt} resides, first, in the infinite number of quadratic optimization problems that we have to solve. In a second time, the quadratic optimization problems are non-necessarily concave and thus can be difficult to solve. 
\section{Results}
\label{mainresults}
This section contains the main contribution of the paper i.e. the computation of the optimal value of Problem~\eqref{pbopt}. This latter computation uses a discretization of initial values set and the computation of a finite integer after which we know that the optimal cannot be reached. In this section, we begin to recall some basic results that we need in our development. We then study the case of linear systems i.e. where $b=0$ in Eq.~\eqref{context}. Finally, we show how to use the linear case to solve the case where the system is affine.   
\subsection{Basic notions}

@@

To complete the computations, we have to deal with $\lfrak$. We suppose that the integer $\ks$ is given and then to compute $\lfrak$ we employ the same technique used for $\mu(\cdot)$. It suffices to browse the vertices of $\xin$ once since we evaluate, for each vertice, $\min\{x^\intercal Q x, x^\intercal {A^{\ks}}^\intercal Q {A^{\ks}}x+q^\intercal A^{\ks} x\}$ and take the maximal value. 
\section{Experiments}
\label{experiments}
We decompose the experiments into two sub-classes: the linear and affine systems. Those sub-classes are separeted in three parts : the  case where the objective function is a quadratic homogeneous function; the linear case and the non-homogeneous quadratic objective function. For each cut, we provide a numerical example.  
\subsection{Linear systems}

@@

\subsubsection{Random matrices}
\section{Conclusion and Future Works}
\label{conclusion}
In this paper, we introduce a new class of optimization problems. This class solves verification problem coming from computer science. This class can be described as follows : the objective function is convex and quadratic and the decision variable is constrained to belong to the reachable values set of a discrete-time affine system which is not a convex set. The main issue is the infinite horizon problem. To maximize the function we can wait for an arbitrary long time. Thus the paper proposes to compute, in the case of stable affine system, a finite horizon after which the search of an optimal value is useless. The problem becomes a finite horizon problem which can be solved in a finite time.
